Dynamical modeling of tidal streams

### 2013-12-27 tex/ms.tex
streams: tweak

Thus, in action--angle coordinates, the dynamics of tidal-stream
formation is simplified to linear growth of small initial angle
differences due to small action/frequency differences. The angle
direction $\unitvec{\veco} \equiv \Delta \veco / | \Delta \veco|$
along which a star moves away from the progenitor is related to the
initial action offset by $\Delta \veco \approx (\partial^2 H /
\partial \vecj \partial \vecj) \Delta \vecj$, in the limit of small
action offsets, because $\veco = \partial H / \partial \vecj$. For an
approximately one-dimensional stream to form, the Hessian matrix
$(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$ has
to be dominated by a single large eigenvalue; in this case the angle
difference of stars stripped from the stream at any time will fall
along approximately the same direction $\unitvec\veco$, and the stream
will essentially be one dimensional. In detail, because the
distribution of initial offsets $\Delta \vecj$ is not isotropic, the
direction $\unitvec \veco$ is not just determined by the potential
(through $H$) and the progenitor orbit, but also by the distribution
of $\Delta \vecj$, which is set through a combination of the internal
dynamics of the progenitor and the stripping process (see discussion
in \citealt{Sanders13b}). For dynamical modeling we can be agnostic
about the relative importance of all of these contributors to
$\unitvec \veco$ by flexible modeling of the distribution of $\Delta
\vecj$ or $\Delta \veco$.

@@

The distribution of frequencies is three dimensional. Because we model
the leading and the trailing arm separately, a single-peaked
distribution suffices to describe the frequency distribution. A
general description in terms of a Gaussian distribution would have
nine free parameters. However, we can model it with fewer free
parameters as follows. First, following \citet{Eyre11a}, we
approximate the \emph{action} distribution as a Gaussian with standard
deviations
\begin{equation}
  \sigma_{J_R} = \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\quad 
  \sigma_{L_Z} = \sigv\,r_{\mathrm{peri}}\,,\quad 
  \sigma_{J_Z} = \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
\end{equation}
where $\sigv$ is the velocity dispersion of the progenitor,
$r_{\mathrm{apo}}$ and $r_{\mathrm{peri}}$ are the apo- and pericenter
radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
the maximum $Z$ reached on this orbit. We assume that the all
correlations between the actions are zero (see
\figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
variance to frequency space using the Hessian $(\partial^2 H /
\partial \vecj \partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian
  $(\partial^2 H / \partial \vecj \partial \vecj)\large|_{\vecj^p}$
  can be computed using the method in \appendixname~\ref{sec:aa} by
  computing $\partial (\vecj,\veca) / \partial (\vecx,\vecv)$ and
  $\partial (\veco,\veca) / \partial (\vecx,\vecv)$ and forming
  $\partial \veco / \partial \vecj = \left(\partial (\vecj,\veca) /
  \partial (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) /
  \partial (\vecx,\vecv)\right)$.}, which gives the variance matrix
$V_\veco$. The principal eigenvector of this matrix is the model
direction along which the stream spreads. While the model I propose
here is a simple one, it works well without having been tweaked for
the simulation under scrutiny. The angle between the principal
eigenvector of $V_\veco$ and the progenitor's frequency vector is
$0.50^\circ$ while the value measured from the simulation is
$0.18^\circ$. For a model with an isotropic action distribution, the
misalignment would be $1.28^circ$.

>>>>

Thus, in action--angle coordinates, the dynamics of tidal-stream
formation is simplified to linear growth of small initial angle
differences due to small action/frequency differences. The angle
direction $\unitvec{\veco} \equiv \Delta \veco / | \Delta \veco|$
along which a star moves away from the progenitor is related to the
initial action offset by $\Delta \veco \approx (\partial \veco /
\partial \vecj) \Delta \vecj = (\partial^2 H / \partial \vecj \partial
\vecj) \Delta \vecj$, in the limit of small action offsets, because
$\veco = \partial H / \partial \vecj$. For an approximately
one-dimensional stream to form, the Hessian matrix $(\partial^2 H /
\partial \vecj \partial \vecj)\large|_{\vecj^p}$ has to be dominated
by a single large eigenvalue; in this case the angle difference of
stars stripped from the stream at any time will fall along
approximately the same direction $\unitvec\veco$, and the stream will
essentially be one dimensional. In detail, because the distribution of
initial offsets $\Delta \vecj$ is not isotropic, the direction
$\unitvec \veco$ is not just determined by the potential (through $H$)
and the progenitor orbit, but also by the distribution of $\Delta
\vecj$, which is set through a combination of the internal dynamics of
the progenitor and the stripping process (see discussion in
\citealt{Sanders13b}). For dynamical modeling we can be agnostic about
the relative importance of all of these contributors to $\unitvec
\veco$ by flexible modeling of the distribution of $\Delta \vecj$ or
$\Delta \veco$.

@@

The distribution of frequencies is three dimensional. Because we model
the leading and the trailing arm separately, a single-peaked
distribution suffices to describe the frequency distribution. A
general description in terms of a Gaussian distribution would have
nine free parameters. However, we can model it with fewer free
parameters as follows. First, following \citet{Eyre11a}, we
approximate the \emph{action} distribution as a Gaussian with standard
deviations
\begin{equation}
  \sigma_{J_R} = \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\quad 
  \sigma_{L_Z} = \sigv\,r_{\mathrm{peri}}\,,\quad 
  \sigma_{J_Z} = \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
\end{equation}
where $\sigv$ is the velocity dispersion of the progenitor,
$r_{\mathrm{apo}}$ and $r_{\mathrm{peri}}$ are the apo- and pericenter
radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
the maximum $Z$ reached on this orbit. We assume that the all
correlations between the actions are zero (see
\figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
variance to frequency space using the Hessian $(\partial \veco /
\partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian $(\partial
  \veco / \partial \vecj)\large|_{\vecj^p}$ can be computed using the
  method in \appendixname~\ref{sec:aa} by computing $\partial
  (\vecj,\veca) / \partial (\vecx,\vecv)$ and $\partial (\veco,\veca)
  / \partial (\vecx,\vecv)$ and forming $\partial \veco / \partial
  \vecj = \left(\partial (\vecj,\veca) / \partial
  (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) / \partial
  (\vecx,\vecv)\right)$.}, which gives the variance matrix
$V_\veco$. The principal eigenvector of this matrix is the model
direction along which the stream spreads. While the model I propose
here is a simple one, it works well without having been tweaked for
the simulation under scrutiny. The angle between the principal
eigenvector of $V_\veco$ and the progenitor's frequency vector is
$0.50^\circ$ while the value measured from the simulation is
$0.18^\circ$. For a model with an isotropic action distribution, the
misalignment would be $1.28^circ$.
